---
title: "Test"
output: html_document
date: "2026-06-15"
editor_options: 
  chunk_output_type: console
---

This is testing a fix that I made for the tree measurements plots.
```{r}
  
  a <- tree_measurements %>% filter(Site == "Brewers Bay" & DBH_cm >= 0 & Mortality != "NA") %>%  
    mutate(Mortality = if_else(Mortality == "Dying", "Alive", Mortality)) %>% 
    group_by(SY, DBH_cm, Mortality) %>% 
    summarise(n = n(), .groups = "drop") %>%
    group_by(SY) %>% 
    mutate(freq = n / sum(n)) %>% 
    ungroup()
  
  
  bins <- seq(2.5, max(a$DBH_cm + 3), by = 3)
  
  
  b <- a %>% 
    group_by(SY, Mortality) %>% 
    nest() %>% 
    mutate(LF = map(data, ~ .x %>%
                      data.frame() %>% 
                      mutate(Bin = cut(DBH_cm, breaks = bins, right = FALSE)) %>%
                      group_by(Bin) %>%
                      summarise(total_freq = sum(freq, na.rm = TRUE)))) %>% 
    unnest(LF) %>% 
    ungroup()
  
  
  br <- unique(b$Bin)
  la <- labeler(bin_num = length(unique(b$Bin)) , bin_size = bin_size)
  
  ggplot(b, aes(x=Bin, y=total_freq, fill=Mortality)) +
    geom_bar(stat="identity", position = "stack", width = .9, color="black", linewidth=.5) +
    scale_x_discrete(labels = c("2.5-3","3-6","6-9","9-12","12-15","15-18","18-21","21-24","24-27","27-30"),
                     limits = factor(br)) +
    scale_fill_manual(values = c("Alive" = "darkolivegreen3", "Dead" = "burlywood4")) +
    xlab(label = "DBH Size Class (cm)") + 
    ylab(label = "Relative Frequency") + 
    ggtitle("Title") +
    theme_Publication() +
    facet_wrap(~SY)



```


Woody Debris Figure

Krauss, K. W., Doyle, T. W., Twilley, R. R., Smith III, T. J., Whelan, K. R., & Sullivan, J. K. (2005). Woody debris in the mangrove forests of South Florida 1. Biotropica: the Journal of Biology and Conservation, 37(1), 9-15.

ALLEN, J. A., K. C. EWEL,B.D.KEELAND,T.TARA, AND T. J. SMITH III. 2000.
Downed wood in Micronesian mangrove forests. Wetlands 20: 169–176.


v=pie2 x sum d2/8L
Mass was calculated using conversions of 0.5 tons m3 for
wood 7.5cm and 0.5,0.35, and 0.20 tons m3 for
larger wood in the sound, intermediate, and rotten categories, respectively.


Our methods

<1cm (2-4m)
  assume diameter 0.5
1-7.5cm (2-6m)
  assume diameter 4.25cm
greater than 7.5 (20m or transect length upon meeting upland veg)
  have diameter for each one


1. Look at counts of fine, intermediate and coarse across sites (Plotting first all plots, then average across sites)

```{r}
library(readr)
library(plotrix)
library(tidyverse)
library(ggforce)
library(patchwork)

Coarse_wood <- read_csv("Data-Kaitlin/QA_QC TMMP Data 6-25-25 - Coarse woody debris.csv")

Wood <- read_csv("Data-Kaitlin/QA_QC TMMP Data 6-25-25 - Wood debris.csv")


#QA/QC Cleaning 

Coarsecheck<-Coarse_wood%>%
  filter(Decay_class=="T")
 
unique(Coarse_wood$Decay_class)
#[1] "I"    "S"    "R"    "T"    "R, T" "I, T" "I, R"

#in the cases where there are two classes listed, selecting the first class, filtering out the single instance of T only 
Coarse_wood<-Coarse_wood%>%
  filter(Decay_class!="T")%>%
  mutate(Decay_class=recode(Decay_class,
                            "R, T"="R",
                            "I, T"="I",
                            "I, R"="I"))

unique(Coarse_wood$Decay_class)
#looks good now! 

unique(Coarse_wood$SY)
unique(Coarse_wood$Island)
unique(Coarse_wood$Site)
unique(Coarse_wood$Plot)
unique(Coarse_wood$Transect_num)


#Will not have coarse debris data for every plot because if there is no coarse debris found then dbh is not entered

#will need to use actual transect length for these calculations, think I need to pull that from fine wood.

Coarse_wood$Decay_class<-factor(Coarse_wood$Decay_class,levels=c("R","I","S"))

#want a count of all the total number of each in each debris class for each plot
CoarseCount<-Coarse_wood%>%
  group_by(SY, Site, Plot, Decay_class)%>%
  summarise(N=n())%>%
  pivot_wider(names_from = Decay_class, values_from =N )



#stacked bar-plot
Cwoodcount<-Coarse_wood%>%
  group_by(SY,Site,Plot,Decay_class)%>%
  summarise(N=n())%>%
  mutate(Site_Plot = paste(Site, Plot, sep = " "))%>%
  ggplot(aes(x=Site_Plot, y=N, fill=Decay_class))+
    geom_col(position="stack")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  facet_wrap(~SY, ncol=1)+
 scale_fill_manual(values=c("#AAC041","#889A34","#444D1B"),labels=c("Rotton","Intermediate","Sound"))+
  ggtitle("Count of Coarse Debris Encountered")+
  xlab("Site")+
  ylab("Woody Debris Count")
  
  Cwoodcount
  
#need to make a plot with counts of coarse on positive axes and fine and medium on negative axes, need to merge coarse counts with fine/medium counts

FineCount<-Wood%>%
    select(SY,Island,Site,Plot,Transect_num,Transect_m, Wood_fine_num, Wood_med_num)

#selecting coarse pivioting wider so I can merge coarse and fine together
Cwoodcount<-Coarse_wood%>%
  group_by(SY,Site,Plot,Transect_num,Decay_class)%>%
  summarise(N=n())%>%
  pivot_wider(names_from =Decay_class , values_from = N)
   
#join the two df together  
CountMerge<-FineCount%>%
  left_join(Cwoodcount, by = c("SY", "Site","Plot", "Transect_num"))

#make fine and medium negative then pivot longer for plotting
CountMerge<-CountMerge%>%
  mutate(Fine=Wood_fine_num*-1, Medium=Wood_med_num*-1)%>%
  pivot_longer(cols=I:Medium, names_to = "Category" , values_to = "Count")

CountMergeSite<-CountMerge%>%
  mutate(Site_Plot = paste(Site, Plot, sep = " "))%>%
  group_by(SY,Site_Plot, Category)%>%
  summarise(Count=sum(Count))



CountMergeSite[is.na(CountMergeSite)] <- 0

CountMergeSite$Category<-factor(CountMergeSite$Category, levels=c("S", "I", "R", "Medium", "Fine"))

#and now we plot!

CountPlot<-CountMergeSite%>%
  ggplot(aes(x=Site_Plot, y=Count, fill=Category))+
    geom_col(position="stack")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  facet_wrap(~SY, ncol=1)+
 scale_fill_manual(values=c("#444D1B","#889A34","#AAC041","#834200","#B5651D"),labels=c("Sound","Intermediate","Rotton", "Medium","Fine"))+
  ggtitle("Total Woody Debris Count Per Plot")+
  xlab("Plot")+
  ylab("Woody Debris Count")

CountPlot


#I want to plot stacked bar with medium and fine on the positive axes stacked per plot

FinePlotStack<-CountMergeSite%>%
  filter(Category=="Fine"|Category=="Medium")%>%
  mutate(Count=Count*-1)%>%
  ggplot(aes(x=Site_Plot, y=Count, fill=Category))+
    geom_col(position="stack")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  facet_wrap(~SY, ncol=1)+
 scale_fill_manual(values=c("#834200","#B5651D"))+
  ggtitle("Count of Fine Woody Debris Count Per Plot")+
  xlab("Plot")+
  ylab("Woody Debris Count")
  
 FinePlotStack 
#point plot showing total counts of coarse debris dot for each plot (ABC)
  
  Cwoodcount<-Coarse_wood%>%
  group_by(SY,Site,Plot)%>%
  summarise(N=n())%>%
  ggplot(aes(x=Site, y=N, color=Plot))+
    geom_point()+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  facet_wrap(~SY, ncol=1)+
    ggtitle("Total Count of Coarse Woody Debris")
  
  Cwoodcount
  

#change dbh from cm to meters
Coarse_wood<-Coarse_wood%>%
  mutate(DBH_m=DBH_cm/100)

#calculate coarse woody debris volume/unit area separately for each debris degradation class 
#this is using the formula from the paper
totaldebris<-Coarse_wood%>%
  group_by(SY, Site, Plot, Transect_num, Decay_class)%>%
  summarise(sumd2=sum(DBH_m^2, na.rm=TRUE), .groups="drop")%>% #adding the squared DBH for each site plot combo
  pivot_wider(names_from = Decay_class , values_from = sumd2)
  
#merging coarse woody debris information with other, sele
Woodmerge<-Wood%>%
  select(SY,Island,Site,Plot,Transect_num,Transect_m, Wood_fine_num, Wood_med_num)

#now i have total diameter of debris for each transect, even if that coarse debris count is zero  
CoarseMerge<-Woodmerge%>%
  left_join(totaldebris, by = c("SY", "Site","Plot", "Transect_num"))

CoarseMerge[is.na(CoarseMerge)] <- 0


#now mutate to get volume for each decay class (units=m3 woody debris/m2), then I multiply by conversion factors to determine mass for decay class) 
CoarseMergeVolume<-CoarseMerge%>%
  mutate(IVol=(pi^2 *I/ (8 * Transect_m)), #m3/m2
         SVol=(pi^2*S/ (8 * Transect_m)), #m3/m2
         RVol=(pi^2*R/(8*Transect_m)),     #m3/m2
               IVol_ha=IVol*10000, #m3/ha  Volume-m3/ha
               SVol_ha=SVol*10000, #m3/ha
               RVol_ha=RVol*10000, #m3/ha
               BMI=IVol_ha*0.35, #t/ha
               BMS=SVol_ha*0.5,  #t/ha
               BMR=RVol_ha*0.2,  #t/ha
         CoarseMass=BMI+BMS+BMR, #total t/ha
         CoarseVol=IVol_ha+SVol_ha+RVol_ha) #total m3/ha


#plotting total count of fine across sites
CoarseMergeVolume$SY<-as.factor(CoarseMergeVolume$SY)

FineCountPlot<-CoarseMergeVolume%>%
  ggplot(aes(x=Site, y=Wood_fine_num, fill =SY))+
  geom_boxplot()+ 
    geom_point(size=1, position=position_dodge(width=0.75))+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  ggtitle("Count of Fine Woody Debris")+
  scale_fill_manual(values=c("#889A34","#B5651D"))+
  theme(axis.text.x = element_blank())

FineCountPlot




#same deal for medium
MedCountPlot<-CoarseMergeVolume%>%
  ggplot(aes(x=Site, y=Wood_med_num, fill =SY))+
  geom_boxplot()+ 
    geom_point(size=1, position=position_dodge(width=0.75))+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  ggtitle("Count of Medium Woody Debris")+
  scale_fill_manual(values=c("#889A34","#B5651D"))

MedCountPlot


SmallCount<-FineCountPlot/MedCountPlot+plot_layout(axis_titles="collect")
  
SmallCount

#point plot total mass of woody debris across sites
MassPlot<-CoarseMergeVolume%>%
  ggplot(aes(x=Site, y=CoarseMass, color=Plot))+
  geom_jitter(width=0.2)+
  facet_wrap(~SY, ncol=1)+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  ggtitle("Mass of Woody Debris")+
  ylab("Coarse Woody Debris (tonnes/ha)")

MassPlot

#calculate mass of fine and medium debris
    #<1cm (2-4m) assume diameter 0.005m (0.5cm)
    #1-7.5cm (2-6m) assume diameter 0.0425 (4.25cm)

#need to determine the number of meters sampled,if transect lengths less than 4 meters for fine and 6 meters for coarse then subtract two from them, otherwise use 2m for fine and 4m for coarse
SmallMedCoarseVolume<-CoarseMergeVolume%>%
  mutate(Fine_m=if_else(Transect_m<4, Transect_m-2, 2))%>% #fine debris 2-4m (if shorter than 4 want true length, else 2m)
  mutate(Med_m=if_else(Transect_m<6, Transect_m-2,4))%>% #medium debris 2-6m
  mutate(FVol=(pi^2*((0.005^2)*Wood_fine_num)/ (8 * Fine_m))*10000, #m2/ha
         MVol=(pi^2*((0.0425^2)*Wood_med_num)/(8*Med_m))*10000, #m2/ha
         FMass=FVol*0.5,#t/ha
         MMass=MVol*0.5, #t/ha
         TotalMass=FMass+MMass+CoarseMass,#t/ha
         TotalVolume=FVol+MVol+CoarseVol)#m2/ha
  


#Total Woody Debris Mass
SmallMedCoarseVolume$SY<-as.factor(SmallMedCoarseVolume$SY)

TotalMassPlot<-SmallMedCoarseVolume%>%
  ggplot(aes(x=Site, y=TotalMass,fill=SY))+
 geom_boxplot()+
  geom_point(size=1, position=position_dodge(width=0.75))+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  ggtitle("Mass of Woody Debris")+
  ylab("Woody Debris (tonnes/ha)")+
  scale_fill_manual(values=c("#889A34","#B5651D"))
TotalMassPlot

#Total Woody Debris Volume
SmallMedCoarseVolume$SY<-as.factor(SmallMedCoarseVolume$SY)

TotalVolumePlot<-SmallMedCoarseVolume%>%
  ggplot(aes(x=Site, y=TotalVolume,fill=SY))+
 geom_boxplot()+
   geom_point(size=1, position=position_dodge(width=0.75))+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  ggtitle("Volume of Woody Debris")+
  ylab("Woody Debris (m3/ha)")+
  scale_fill_manual(values=c("#889A34","#B5651D"))
TotalVolumePlot
  
#mean and SEM plot
MeanVolPlot<-SmallMedCoarseVolume%>%
  mutate(SmallVol=FVol+MVol)%>%
  group_by(SY, Site)%>%
  summarise(MeanCoarse=mean(CoarseVol), MeanFine=-mean(SmallVol), SEMCoarse=std.error(CoarseVol), SEMFine=std.error(SmallVol),.groups = "drop")%>%
  pivot_longer(
    cols = c(MeanCoarse, MeanFine, SEMCoarse, SEMFine),
    names_to = c(".value", "type"),
    names_pattern = "(Mean|SEM)(Coarse|Fine)"
  )%>%
  ggplot(aes(x=Site, y=Mean, fill=type))+
  geom_col(stat="identity")+
    geom_errorbar(aes(
    ymin = Mean - SEM,
    ymax = Mean + SEM), 
    width=0.2, 
    position = position_identity()
  ) +
  geom_hline(yintercept = 0)+
  facet_wrap(~SY, ncol=1)+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  scale_fill_manual(values=c("#889A34","#B5651D"))

MeanVolPlot


#plot with coarse volume and fine and medium counts
#want to do with stacked bars, can do this per transect, but not sure how to look at data with means per site 

VolumePosCountNeg<-SmallMedCoarseVolume%>%
  mutate(PlotWoodFine=-1*Wood_fine_num,
         PlotWoodMedium=-1*Wood_med_num)%>%
  select(SY,Island, Site,Plot,Transect_num,PlotWoodFine, PlotWoodMedium, IVol_ha,SVol_ha,RVol_ha)%>%
  pivot_longer(cols=c(PlotWoodFine,PlotWoodMedium, IVol_ha, SVol_ha, RVol_ha), names_to = "Category", values_to = "Total")%>%
  filter(Site=="Brewers Bay")%>%
  ggplot(aes(x=Transect_num,y=Total, fill=Category))+
           geom_col(position="stack")

VolumePosCountNeg
#I dont like this plot path situation thing happening here, taking a pause on this situation 

```

Want to cleanly pull all the info that I mathed out of the woody debris info for each site so I can plot it.Maybe mean for each site?? Will rename columns including units in column names for ease later

Coarse S Volume, Mass 
Coarse I Volume, Mass 
Coarse R Volume, Mass
Total Coarse Count, Mass, and Volume
Fine Mass, Volume
Medium Mass, Volume
```{r}

WoodyDebrisSummaryMeanSEM<-SmallMedCoarseVolume%>%
  select(SY,Island,Site,Plot,Transect_num,Transect_m,Wood_fine_num,Wood_med_num, SVol_ha,BMS,IVol_ha,BMI, RVol_ha, BMR, CoarseMass, CoarseVol, FVol, MVol, FMass, MMass,TotalMass, TotalVolume)%>%
  group_by(SY, Island, Site)%>%
  summarise(FineCount_n=mean(Wood_fine_num),
            SEMFineCount=std.error(Wood_fine_num),
            MedCount_n=mean(Wood_med_num),
            SemMedCount=std.error(Wood_med_num),
            FineVol_m2ha=mean(FVol),
            SEMFineVol=std.error(FVol),
            MedVol_m2ha=mean(MVol),
            SEMMedVol=std.error(MVol),
            FineMass_tha=mean(FMass),
            SEMFineMass=std.error(FMass),
            MedMass_tha=mean(MMass),
            SEMMedMass=std.error(MMass),
            SoundVol_m2ha=mean(SVol_ha),
            SEMSoundVol=std.error(SVol_ha),
            SoundMass_tha=mean(BMS),
            SEMSoundMass=std.error(BMS),
            InterVol_m2ha=mean(IVol_ha),
            SEMIVol=std.error(IVol_ha),
            InterMass_tha=mean(BMI),
            SEMIMass=std.error(BMI),
            RotVol_m2ha=mean(RVol_ha),
            SEMRVol=std.error(RVol_ha),
            RotMass_tha=mean(BMR),
            SEMRMass=std.error(BMR),
            CoarseMass_tha=mean(CoarseMass),
            SEMCoarseMass=std.error(CoarseMass),
            CoarseVol_m2ha=mean(CoarseVol),
            SEMCoarseVol=std.error(CoarseVol),
            TotalMass_tha=mean(TotalMass),
            SEMTotalMass_tha=std.error(TotalMass),
            TotalVolume_m2ha=mean(TotalVolume),
            SEMTotalVolume=std.error(TotalVolume)
            )


```

```{r}
library(readr)
library(plotrix)
library(tidyverse)
library(ggforce)

Coarse_wood <- read_csv("Data-Kaitlin/QA_QC TMMP Data 6-25-25 - Coarse woody debris.csv")

Wood <- read_csv("Data-Kaitlin/QA_QC TMMP Data 6-25-25 - Wood debris.csv")


#QA/QC Cleaning 

Coarsecheck<-Coarse_wood%>%
  filter(Decay_class=="T")
 
unique(Coarse_wood$Decay_class)
#[1] "I"    "S"    "R"    "T"    "R, T" "I, T" "I, R"

#in the cases where there are two classes listed, selecting the first class, filtering out the single instance of T only 
Coarse_wood<-Coarse_wood%>%
  filter(Decay_class!="T")%>%
  mutate(Decay_class=recode(Decay_class,
                            "R, T"="R",
                            "I, T"="I",
                            "I, R"="I"))

unique(Coarse_wood$Decay_class)
#looks good now! 

unique(Coarse_wood$SY)
unique(Coarse_wood$Island)
unique(Coarse_wood$Site)
unique(Coarse_wood$Plot)
unique(Coarse_wood$Transect_num)


#Will not have coarse debris data for every plot because if there is no coarse debris found then dbh is not entered

#will need to use actual transect length for these calculations, think I need to pull that from fine wood.

Coarse_wood$Decay_class<-factor(Coarse_wood$Decay_class,levels=c("R","I","S"))

#want a count of all the total number of each in each debris class for each plot
CoarseCount<-Coarse_wood%>%
  group_by(SY, Site, Plot, Decay_class)%>%
  summarise(N=n())%>%
  pivot_wider(names_from = Decay_class, values_from =N )



#stacked bar-plot
Cwoodcount<-Coarse_wood%>%
  group_by(SY,Site,Plot,Decay_class)%>%
  summarise(N=n())%>%
  mutate(Site_Plot = paste(Site, Plot, sep = " "))%>%
  ggplot(aes(x=Site_Plot, y=N, fill=Decay_class))+
    geom_col(position="stack")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  facet_wrap(~SY, ncol=1)+
 scale_fill_manual(values=c("#AAC041","#889A34","#444D1B"),labels=c("Rotton","Intermediate","Sound"))+
  ggtitle("Count of Coarse Debris Encountered")+
  xlab("Site")+
  ylab("Woody Debris Count")
  
  Cwoodcount
  
#need to make a plot with counts of coarse on positive axes and fine and medium on negative axes, need to merge coarse counts with fine/medium counts

FineCount<-Wood%>%
    select(SY,Island,Site,Plot,Transect_num,Transect_m, Wood_fine_num, Wood_med_num)

#selecting coarse pivioting wider so I can merge coarse and fine together
Cwoodcount<-Coarse_wood%>%
  group_by(SY,Site,Plot,Transect_num,Decay_class)%>%
  summarise(N=n())%>%
  pivot_wider(names_from =Decay_class , values_from = N)
   
#join the two df together  
CountMerge<-FineCount%>%
  left_join(Cwoodcount, by = c("SY", "Site","Plot", "Transect_num"))

#make fine and medium negative then pivot longer for plotting
CountMerge<-CountMerge%>%
  mutate(Fine=Wood_fine_num*-1, Medium=Wood_med_num*-1)%>%
  pivot_longer(cols=I:Medium, names_to = "Category" , values_to = "Count")

CountMergeSite<-CountMerge%>%
  mutate(Site_Plot = paste(Site, Plot, sep = " "))%>%
  group_by(SY,Site_Plot, Category)%>%
  summarise(Count=sum(Count))



CountMergeSite[is.na(CountMergeSite)] <- 0

CountMergeSite$Category<-factor(CountMergeSite$Category, levels=c("S", "I", "R", "Medium", "Fine"))

#and now we plot!

CountPlot<-CountMergeSite%>%
  ggplot(aes(x=Site_Plot, y=Count, fill=Category))+
    geom_col(position="stack")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  facet_wrap(~SY, ncol=1)+
 scale_fill_manual(values=c("#444D1B","#889A34","#AAC041","#834200","#B5651D"),labels=c("Sound","Intermediate","Rotton", "Medium","Fine"))+
  ggtitle("Total Woody Debris Count Per Plot")+
  xlab("Plot")+
  ylab("Woody Debris Count")

CountPlot


#I want to plot stacked bar with medium and fine on the positive axes stacked per plot

FinePlotStack<-CountMergeSite%>%
  filter(Category=="Fine"|Category=="Medium")%>%
  mutate(Count=Count*-1)%>%
  ggplot(aes(x=Site_Plot, y=Count, fill=Category))+
    geom_col(position="stack")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  facet_wrap(~SY, ncol=1)+
 scale_fill_manual(values=c("#834200","#B5651D"))+
  ggtitle("Count of Fine Woody Debris Count Per Plot")+
  xlab("Plot")+
  ylab("Woody Debris Count")
```



Need to calculate total basal area for each plot
```{r}
library(readr)
library(tidyverse)

radial<- read_csv("Data-Kaitlin/QA_QC TMMP Data 6-25-25 - Tree measurements.csv")
#plan is to first convert DBH's to basal areas, need it to be in m2/ha

#cm->m divide by 100
#dbh/2 to get radius, then pi*radius^2
#then sum total for plot
#divide by plot area 5.65m radius (pi * 5.62^2), not sure if I want to do this before or after I combine at the site level. prolly before cuz all my woody debris is at the plot level 

#then I'll need to combine this info with the woody debris info to make the plot that I want



```

I think Ideally the summarized data should be combined into a single data frame, so different metrics can be plotted against each other
```{r}



```

